\Title

### 2023-02-07 Grimaldi-etal-BiolCybernetics.tex
discussion : c

Such recurrent activity would be essential for the implementation of predictive or anticipatory processes. This is essential in a neural system because it contains several different delays that require temporal alignment~\citep{hogendoorn_predictive_2019}. This has been modeled before to explain, for example, the flash-lag illusion~\citep{khoei_flash-lag_2017}. %\note{Say that this could be implemented by properly setting the timing of the label} 
As mentioned previously, this could be implemented using generalized coordinates (i.e., variables such as position complemented by velocity, acceleration, jerk, \ldots), and ``neurobiologically, using delay operators just means changing synaptic connection strengths to take different mixtures of generalized sensations and their prediction errors''~\citep{perrinet_active_2014}. Our proposed model using heterogeneous delays provides an alternative and elegant implementation solution to this problem.
%
\subsection{Perspectives}
%%%-----------------------------------------------------------------
% \note{la tache ne met pas en avant les capacites des motifs hetero synaptiques =trop simple et peut se faire en HOTS}
In the definition of our task, we highlighted how the generation of events depends on the spatial gradient in each image. This gradient is defined in both horizontal and vertical dimensions and its maxima are generally oriented. Taken together, these oriented edges form the contours of visual objects in the scene~\citep{koenderink_representation_1987}. Thus, there is a dependence within the event stream between motion and orientation information. It would be crucial to study this interdependence further. This could be initiated by training the model on a dataset with labels providing local orientation. %\note{we can say we tried and that it works well}. 
Studying this dependence will allow us to dissociate these two forms of visual information and to enable us to integrate them. In particular, it will allow us to consider that the definition of motion is more accurate perpendicular to an oriented contour (aka, the aperture problem), thus allowing us to implement recurrent predictive rules, such as those identified to dissociate this problem~\citep{perrinet_motion-based_2012}.

The model is trained on a low-level local motion detection task, and one might wonder if it could be trained on higher-level tasks. An example task would be depth estimation in the visual scene. There are multiple sources of information for inferring depth, such as binocular disparity or changes in texture or shading, but in our case, motion parallax would be the most significant cue~\citep{rogers_motion_1979}. This is because objects close to an observer move relatively faster on the retina than an object at a great distance, and also that visual occlusions depend on depth ordering. Using this information, one can segment objects and estimate their depth~\citep{yoonessi_contribution_2011}. However, this would require computing first the optical flow, that is extending the framework described here for a rigid full-field motion to a more generic one where motion may vary in the visual field. A possible implementation is therefore to add a new layer to our model, in analogy with the hierarchical organization highlighted in the visual cortex. This is theoretically possible by using the output of our model (which estimates velocity in retinotopic space) as input to a new layer of neurons that would estimate velocity in the visual field, including the depth dimension in the output supervision labels. This could have direct and important applications, for example in autonomous driving, to detect obstacles in a fast and robust way. Another extension would be to actively generate sensor movements (physically or virtually) to yield better depths estimates, notably to disambiguate uncertain estimates~\citep{nawrot_eye_2003}.

In conclusion, the model we have presented provides a way to efficiently process event-based signals. We have shown that we can train the model in a semi-supervised manner, knowing \emph{what} output label but not knowing \emph{when} it occurs. Another perspective would be to extend the model to a fully self-supervised learning paradigm, i.e., without any labeled data~\citep{barlow_unsupervised_1989}. This type of learning is thought to be predominant in the central nervous system and, assuming that the signal is sparse~\citep{olshausen_emergence_1996}, one could extend those Hebbian sparse learning schemes to spikes~\citep{perrinet_emergence_2004, masquelier_competitive_2009}. We anticipate that this would be particularly well suited to the exploration of neurobiological data. In fact, there is a substantial literature indicating that brain dynamics often organize into stereotyped sequences such as synfire chains~\citep{ikegaya_synfire_2004}, packets~\citep{luczak_sequential_2007}, or hippocampal sequences~\citep{pastalkova_internally_2008, villette_internally_2015}. These motifs are stereotypical and robust, as they can be activated in the same motif from day to day~\citep{haimerl_internal_2019}. In contrast to conventional methodologies that are used to process neurobiological data, such a event-based model would be able to answer key questions regarding the representation of information in neurobiological data. Furthermore, it would open up possibilities in the field of machine learning, especially in computer vision, to address current key concerns such as robustness to attacks, scalability, interpretability, or energy consumption.
%
%%%%%%%%%%%%%%%%%%%%%%%%
\backmatter
%
\bmhead{Acknowledgments}
%
% Acknowledgement hidden for review.
\Acknowledgments
%
%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

Such recurrent activity would be essential for the implementation of predictive or anticipatory processes. This is essential in a neural system because it contains several different delays that require temporal alignment~\citep{hogendoorn_predictive_2019}. This has been modeled before to explain, for example, the flash-lag illusion~\citep{khoei_flash-lag_2017}. %\note{Say that this could be implemented by properly setting the timing of the label} 
As mentioned previously, this could be implemented using generalized coordinates (i.e., variables such as position complemented by velocity, acceleration, jerk, \ldots), and ``neurobiologically, using delay operators just means changing synaptic connection strengths to take different mixtures of generalized sensations and their prediction errors''~\citep{perrinet_active_2014}. Our proposed model using heterogeneous delays provides an alternative and elegant implementation solution to this problem.
%
\subsection{Perspectives}
%%%-----------------------------------------------------------------
% \note{la tache ne met pas en avant les capacites des motifs hetero synaptiques =trop simple et peut se faire en HOTS}
In defining our task, we emphasized that the generation of events depends on the spatial gradient in each image. This gradient has both horizontal and vertical dimensions, and its maxima are generally orientation dependent. Taken together, these oriented edges form the contours of visual objects in the scene~\citep{koenderink_representation_1987}. Thus, there is an interdependence between the information about motion and the information about orientation within the event stream. It would be crucial to investigate this dependency further. This could be initiated by training the model on a dataset with labels that provide local orientation. %\note{we can say that we tried it and that it works well}. 
Exploring this dependence will allow us to dissociate these two forms of visual information and enable us to integrate them. In particular, it will allow us to consider that the definition of motion is more accurate perpendicular to an oriented contour (aka the aperture problem). Thus, it will allow us to implement recurrent prediction rules, such as those identified to dissociate this problem~\citep{perrinet_motion-based_2012}.

The model is trained on a low-level local motion detection task, and one might wonder if it could be trained on higher-level tasks. An example of such a task would be the estimation of depth in the visual scene. There are several sources of information for depth estimation, such as binocular disparity or changes in texture or shading, but in our case motion parallax would be the most important cue~\citep{rogers_motion_1979}. This is because objects that are close to an observer will move relatively faster on the retina than an object that is far away, and also because visual occlusions are dependent on the depth order. Using this information, it is possible to segment objects and estimate their depth~\citep{yoonessi_contribution_2011}. However, this would require the computation of the optical flow first, i.e., the extension of the framework described here for a rigid full-field motion to a more general one where the motion may vary in the visual field. A possible implementation is therefore to add a new layer to our model, analogous to the hierarchical organization highlighted in the visual cortex. This is theoretically possible by using the output of our model (which estimates velocity in retinotopic space) as input to a new layer of neurons that would estimate velocity in the visual field, including the depth dimension in the output supervision labels. This could have direct and important applications, e.g. in autonomous driving to detect obstacles in a fast and robust way. Another extension would be to actively generate sensor motion (physical or virtual) to obtain better depth estimates, especially to disambiguate uncertain estimates~\citep{nawrot_eye_2003}.

In conclusion, the model that we have presented provides a way to process event-based signals in an efficient manner. We have shown that we can train the model semi-supervised, knowing \emph{what} output label, but not \emph{when} it occurs. Another perspective would be to extend the model to a fully self-supervised learning paradigm, i.e., without any labeled data~\citep{barlow_unsupervised_1989}. This type of learning is thought to be prevalent in the central nervous system and, assuming the signal is sparse~\citep{olshausen_emergence_1996}, one could extend these Hebbian sparse learning schemes to spikes~\citep{perrinet_emergence_2004, masquelier_competitive_2009}. We expect that this would be particularly useful for exploring neurobiological data. Indeed, there is a large literature showing that brain dynamics often organize into stereotyped sequences, such as synfire chains~\citep{ikegaya_synfire_2004}, packets~\citep{luczak_sequential_2007}, or hippocampal sequences~\citep{pastalkova_internally_2008, villette_internally_2015}. These patterns are stereotyped and robust, as they can be activated in the same pattern from day to day~\citep{haimerl_internal_2019}. In contrast to conventional methods of processing neurobiological data, such an event-based model would be able to answer key questions about the representation of information in neurobiological data, and it would open up possibilities in the field of computational neuroscience. Furthermore, it would open up possibilities in the field of machine learning, especially in computer vision, to address current key concerns such as robustness to attacks, scalability, interpretability, or energy consumption.
%
%%%%%%%%%%%%%%%%%%%%%%%%
\backmatter
%
\bmhead{Acknowledgments}
%
% Acknowledgement hidden for review.
\Acknowledgments
%
%%%%%%%%%%%%%%%%%%%%%%%%
